Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-31 Manuscript/manuscript.Rmd
added refs to manuscript

Identifying the mechanisms underlying the formation of TADs are a complex and active area of research. The distinct patterns of epigenomic marks, binding peaks of different factors, and actively transcribed genes [??? REFs] prompt the use of computational approaches to identify which epigenomic features are most predictive of the TAD boundaries. Current methods rely on the use of classification algorithms to draw associations between the spatial relationship of functional genomic elements and regions that include TAD boundaries [@mourad2016computational; @hong2017computational; @huang2015predicting]. Although several useful insights have been made, such as the convergent binding of CTCF [??? Elaborate with refs.], many methods ignore key characteristics of domain data, such as heavily imbalanced number of boundaries vs. other regions. In a classification setting [??? are we talking about classification or prediction?], when one is attempting to discriminate between regions with and without TAD boundaries, this will contribute to small imbalance ratios (defined as the ratio of minority to majority classes) [@Orriols-Puig2008]. It is commonly agreed upon in the machine learning community that imbalanced datasets adversely impact the performance of the classifiers as the learned model is biased towards the majority class to minimize the overall error rate [@lusa2010class;@chen2013influence;@jeni2013facing]. Similarly, careful consideration must also be taken when deciding which performance metric(s) to use for evaluation of predictive models, especially in the case of imbalanced data.

@@

Replicate Hi-C data for the GM12878, K562, ... cell lines were downloaded from GEO GSE63525 (Experiments HIC001 to HIC018, HIC069 to HIC 074, ... respectively; produced using the MbolI restriction enzyme) [@rao20143d]. [??? Need supplementary ref.] The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the Arrowhead algorithm [@durand2016juicebox]. The start and end coordinates of all TADs were concatenated, sorted, and unique TAD boundary points were obtained.

>>>>

Identifying the mechanisms underlying the formation of TADs are a complex and active area of research. The distinct patterns of epigenomic marks, binding peaks of different factors, and actively transcribed genes [@xiao2012comparative; @milosavljevic2011emerging] [??? REFs; !!!added refs] prompt the use of computational approaches to identify which epigenomic features are most predictive of the TAD boundaries. Current methods rely on the use of classification algorithms to draw associations between the spatial relationship of functional genomic elements and regions that include TAD boundaries [@mourad2016computational; @hong2017computational; @huang2015predicting]. Although several useful insights have been made, such as the convergent binding of CTCF [@ong2014ctcf; @ghirlando2016ctcf] [??? Elaborate with refs.; !!!added refs], many methods ignore key characteristics of domain data, such as heavily imbalanced number of boundaries vs. other regions. In a classification setting [??? are we talking about classification or prediction?; !!!prediction is a byproduct of classification correct?; wait to discuss in meeting], when one is attempting to discriminate between regions with and without TAD boundaries, this will contribute to small imbalance ratios (defined as the ratio of minority to majority classes) [@Orriols-Puig2008]. As has been reported in the machine learning community, imbalanced datasets are known to adversely impact the performance of the classifiers as the learned model is biased towards the majority class to minimize the overall error rate [@lusa2010class;@chen2013influence;@jeni2013facing]. Similarly, careful consideration must also be taken when deciding which performance metric(s) to use for evaluation of predictive models, especially in the case of imbalanced data.

@@

Replicate Hi-C data for the GM12878 and K562 cell lines were downloaded from GEO GSE63525 [@rao20143d] (Supplementary Table 4) [??? Need supplementary ref.; !!!added supp table] The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the Arrowhead algorithm [@durand2016juicebox]. The start and end coordinates of all TADs were concatenated, sorted, and unique TAD boundary points were obtained.
